---
title: "Modeling 3"
output: html_notebook
---

Importing data
```{r}
roadie <- read.csv("D:/Insurance Project/data.csv")

head(roadie)
acci <- roadie #backup
```

Structure of the data
```{r}
str(roadie)
```

Shape of the data
```{r}
dim(roadie)
```

Variables in the data
```{r}
colnames(roadie)
```

Taking a sample to work with
```{r}
roadie <- roadie[sample(nrow(roadie),50000),]
```

Shape of new dataframe
```{r}
head(roadie)
```
Fixing datetime variables 
```{r}
roadie$Start_Time <- as.Date(roadie$Start_Time)
roadie$End_Time <- as.Date(roadie$End_Time)
roadie$Weather_Timestamp = as.Date(roadie$Weather_Timestamp)
```

How many null values are there in the dataframe?
```{r}
sum(is.na(roadie))
```
Percentage of missing values in each column
```{r}
(colSums(is.na(roadie))/50000)*100 
```
Remove number column since it has 69% missing values
```{r}
roadie$Number <- NULL
```

Removing all null values from the dataframe
```{r}
roadie <- roadie[complete.cases(roadie),]
```

Converting character type to factor types
```{r}
roadie$Street <- as.factor(roadie$Street)
roadie$Side <- as.factor(roadie$Side)
roadie$City <- as.factor(roadie$City)
roadie$State <- as.factor(roadie$State)
roadie$Zipcode <- as.integer(roadie$Zipcode)
roadie$Country <- as.factor(roadie$Country)
roadie$Timezone <- as.factor(roadie$Timezone)
roadie$Airport_Code <- as.factor(roadie$Airport_Code)
roadie$Wind_Direction <- as.factor(roadie$Wind_Direction)
roadie$Weather_Condition <-as.factor(roadie$Weather_Condition)
roadie$Amenity <- as.factor(roadie$Amenity)
roadie$Bump <- as.factor(roadie$Bump)
roadie$Crossing <- as.factor(roadie$Crossing)
roadie$Give_Way <- as.factor(roadie$Give_Way)
roadie$Junction <- as.factor(roadie$Junction)
roadie$No_Exit <- as.factor(roadie$No_Exit)
roadie$Railway <- as.factor(roadie$Railway)
roadie$Roundabout <- as.factor(roadie$Roundabout)
roadie$Station <- as.factor(roadie$Station)
roadie$Stop <- as.factor(roadie$Stop)
roadie$Traffic_Calming <- as.factor(roadie$Traffic_Calming)
roadie$Traffic_Signal <- as.factor(roadie$Traffic_Signal)
roadie$Turning_Loop <- as.factor(roadie$Turning_Loop)
roadie$Sunrise_Sunset <- as.factor(roadie$Sunrise_Sunset)
roadie$Civil_Twilight <- as.factor(roadie$Civil_Twilight)
roadie$Nautical_Twilight <- as.factor(roadie$Nautical_Twilight)
roadie$Astronomical_Twilight <- as.factor(roadie$Astronomical_Twilight)
roadie$Severity <- as.factor(roadie$Severity)
```

Checking correlation using Chi-square test between all the variables and Severity
```{r}
chisq.test(roadie$ID,roadie$Severity)
chisq.test(roadie$Severity,roadie$Start_Time)
chisq.test(roadie$Severity,roadie$End_Time)
chisq.test(roadie$Severity,roadie$Start_Lat)
chisq.test(roadie$Severity,roadie$Start_Lng)
chisq.test(roadie$Severity,roadie$End_Lat)
chisq.test(roadie$Severity,roadie$End_Lng)
chisq.test(roadie$Severity,roadie$Distance.mi.)
chisq.test(roadie$Severity,roadie$Description)
chisq.test(roadie$Severity,roadie$Street)
chisq.test(roadie$Severity,roadie$Side)
chisq.test(roadie$Severity,roadie$City)
chisq.test(roadie$Severity,roadie$County)
chisq.test(roadie$Severity,roadie$State)
chisq.test(roadie$Severity,roadie$Zipcode)
#chisq.test(roadie$Severity,roadie$Country)
chisq.test(roadie$Severity,roadie$Timezone)
chisq.test(roadie$Severity,roadie$Airport_Code)
chisq.test(roadie$Severity,roadie$Weather_Timestamp)
chisq.test(roadie$Severity,roadie$Temperature.F.)
chisq.test(roadie$Severity,roadie$Wind_Chill.F.)
chisq.test(roadie$Severity,roadie$Humidity...)
chisq.test(roadie$Severity,roadie$Pressure.in.)
chisq.test(roadie$Severity,roadie$Visibility.mi.)
chisq.test(roadie$Severity,roadie$Wind_Direction)
chisq.test(roadie$Severity,roadie$Wind_Speed.mph.)
chisq.test(roadie$Severity,roadie$Precipitation.in.)
chisq.test(roadie$Severity,roadie$Weather_Condition)
chisq.test(roadie$Severity,roadie$Amenity)
chisq.test(roadie$Severity,roadie$Bump)
chisq.test(roadie$Severity,roadie$Crossing)
chisq.test(roadie$Severity,roadie$Give_Way)
chisq.test(roadie$Severity,roadie$Junction)
chisq.test(roadie$Severity,roadie$No_Exit)
chisq.test(roadie$Severity,roadie$Railway)
chisq.test(roadie$Severity,roadie$Roundabout)
chisq.test(roadie$Severity,roadie$Station)
chisq.test(roadie$Severity,roadie$Stop)
chisq.test(roadie$Severity,roadie$Traffic_Calming)
chisq.test(roadie$Severity,roadie$Traffic_Signal)
#chisq.test(roadie$Severity,roadie$Turning_Loop)
chisq.test(roadie$Severity,roadie$Sunrise_Sunset)
chisq.test(roadie$Severity,roadie$Civil_Twilight)
chisq.test(roadie$Severity,roadie$Nautical_Twilight)
chisq.test(roadie$Severity,roadie$Astronomical_Twilight)
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

