---
title: "Modeling"
output: html_notebook
---

Traffic Attributes (12):
ID: This is a unique identifier of the accident record.
Source: Indicates source of the accident report (i.e. the API which reported the accident.).
TMC: A traffic accident may have a Traffic Message Channel (TMC) code which provides more detailed description of the event.
Severity: Shows the severity of the accident, a number between 1 and 4, where 1 indicates the least impact on traffic (i.e., short delay as a result of the accident) and 4 indicates a significant impact on traffic (i.e., long delay).
Start_Time: Shows start time of the accident in local time zone.
End_Time: Shows end time of the accident in local time zone.
Start_Lat: Shows latitude in GPS coordinate of the start point.
Start_Lng: Shows longitude in GPS coordinate of the start point.
End_Lat: Shows latitude in GPS coordinate of the end point.
End_Lng: Shows longitude in GPS coordinate of the end point.
Distance(mi): The length of the road extent affected by the accident.
Description: Shows natural language description of the accident.

Address Attributes (9):
Number: Shows the street number in address field.
Street: Shows the street name in address field.
Side: Shows the relative side of the street (Right/Left) in address field.
City: Shows the city in address field.
County: Shows the county in address field.
State: Shows the state in address field.
Zipcode: Shows the zipcode in address field.
Country: Shows the country in address field.
Timezone: Shows timezone based on the location of the accident (eastern, central, etc.).

Weather Attributes (11):
Airport_Code: Denotes an airport-based weather station which is the closest one to location of the accident.
Weather_Timestamp: Shows the time-stamp of weather observation record (in local time).
Temperature(F): Shows the temperature (in Fahrenheit).
Wind_Chill(F): Shows the wind chill (in Fahrenheit).
Humidity(%): Shows the humidity (in percentage).
Pressure(in): Shows the air pressure (in inches).
Visibility(mi): Shows visibility (in miles).
Wind_Direction: Shows wind direction.
Wind_Speed(mph): Shows wind speed (in miles per hour).
Precipitation(in): Shows precipitation amount in inches, if there is any.
Weather_Condition: Shows the weather condition (rain, snow, thunderstorm, fog, etc.).

POI Attributes (13):
Amenity: A Point-Of-Interest (POI) annotation which indicates presence of amenity in a nearby location.
Bump: A POI annotation which indicates presence of speed bump or hump in a nearby location.
Crossing: A POI annotation which indicates presence of crossing in a nearby location.
Give_Way: A POI annotation which indicates presence of give_way sign in a nearby location.
Junction: A POI annotation which indicates presence of junction in a nearby location.
No_Exit: A POI annotation which indicates presence of no_exit sign in a nearby location.
Railway: A POI annotation which indicates presence of railway in a nearby location.
Roundabout: A POI annotation which indicates presence of roundabout in a nearby location.
Station: A POI annotation which indicates presence of station (bus, train, etc.) in a nearby location.
Stop: A POI annotation which indicates presence of stop sign in a nearby location.
Traffic_Calming: A POI annotation which indicates presence of traffic_calming means in a nearby location.
Traffic_Signal: A POI annotation which indicates presence of traffic_signal in a nearby location.
Turning_Loop: A POI annotation which indicates presence of turning_loop in a nearby location.

Period-of-Day (4):
Sunrise_Sunset: Shows the period of day (i.e. day or night) based on sunrise/sunset.
Civil_Twilight: Shows the period of day (i.e. day or night) based on civil twilight.
Nautical_Twilight: Shows the period of day (i.e. day or night) based on nautical twilight.
Astronomical_Twilight: Shows the period of day (i.e. day or night) based on astronomical twilight.

```{r}
library(ggplot2)
library(dplyr)
library(randomForest)
library(caTools)
library(rpart)
library(caret)
library(e1071)
library(ISLR)
library(randomForest)
library(mlr3)
```

```{r}
df <- read.csv("df.csv")
```

```{r}
head(df)
```

```{r}
# fix datetime type
df$Start_Time <- as.Date(df$Start_Time)
df$End_Time <- as.Date(df$End_Time)
df$Weather_Timestamp = as.Date(df$Weather_Timestamp)

```

<!-- Dropping columns that are not relevant -->
<!-- ```{r} -->
<!-- df <- dplyr::select(df,-c(ID,Description,`Distance(mi)`,End_Time,Duration,End_Lat,End_Lng)) -->
<!-- #head(df) -->
<!-- ``` -->

```{r}
df$X = NULL
df$ID = NULL
df$Description = NULL
df$Distance.mi. = NULL
df$End_Time = NULL
df$End_Lat = NULL
df$End_Lng = NULL
df$Number = NULL
```

```{r}
dim(df)
```


```{r}
df$Wind_Chill.F. = NULL #irrelevant

```


```{r}
df <- df[complete.cases(df),]
```

```{r}
summary(df)
```

# sum(is.na(df)) #sum
# (colSums(is.na(df))/1516064)*100 #percentage of missing vals
```{r}
df$Severity <- as.factor(df$Severity)
```


<!-- dropping observations with missing values -->

#df %>% drop_na(City)
#df[complete.cases(df$City),]
#df %>% na.omit(df$City)
#df[complete.cases(df$Zipcode),]
#df %>% drop_na(Zipcode)
#df[complete.cases(df$Airport_Code),]
#df %>% drop_na(Airport_Code)
#df[complete.cases(df$Sunrise_Sunset),]
#df %>% drop_na(Sunrise_Sunset)
#df <- na.omit(df$Sunrise_Sunset)
#df[complete.cases(df$Civil_Twilight),]
#df <- na.omit(df$Civil_Twilight)
#df[complete.cases(df$Nautical_Twilight),]
#df <- na.omit(df$Nautical_Twilight)
#df[complete.cases(df$Astronomical_Twilight),]
#df <- na.omit(df$Astronomical_Twilight)
#df <- df[rowSums(is.na(df)) == 0,]


```{r}
df$County = NULL
```


```{r}
library(tidyr)
```

```{r}
head(df)
```
```{r}
#convert to factors
df$Street <- as.factor(df$Street)
df$Side <- as.factor(df$Side)
df$City <- as.factor(df$City)
df$State <- as.factor(df$State)
df$Zipcode <- as.integer(df$Zipcode)
df$Country <- as.factor(df$Country)
df$Timezone <- as.factor(df$Timezone)
df$Airport_Code <- as.factor(df$Airport_Code)
df$Wind_Direction <- as.factor(df$Wind_Direction)
df$Weather_Condition <-as.factor(df$Weather_Condition)
df$Amenity <- as.factor(df$Amenity)
df$Bump <- as.factor(df$Bump)
df$Crossing <- as.factor(df$Crossing)
df$Give_Way <- as.factor(df$Give_Way)
df$Junction <- as.factor(df$Junction)
df$No_Exit <- as.factor(df$No_Exit)
df$Railway <- as.factor(df$Railway)
df$Roundabout <- as.factor(df$Roundabout)
df$Station <- as.factor(df$Station)
df$Stop <- as.factor(df$Stop)
df$Traffic_Calming <- as.factor(df$Traffic_Calming)
df$Traffic_Signal <- as.factor(df$Traffic_Signal)
df$Turning_Loop <- as.factor(df$Turning_Loop)
df$Sunrise_Sunset <- as.factor(df$Sunrise_Sunset)
df$Civil_Twilight <- as.factor(df$Civil_Twilight)
df$Nautical_Twilight <- as.factor(df$Nautical_Twilight)
df$Astronomical_Twilight <- as.factor(df$Astronomical_Twilight)
```
```{r}
df <- df[complete.cases(df),]
```

```{r}
str(df)
```
```{r}
df$Turning_Loop <- NULL #one factor
df$Country <- NULL
```

```{r}
sev <- df$Severity # save target column to an array before encoding and scaling
```

```{r}
library(caret)

dmy <- dummyVars(" ~.", data=df)
dat_transformed <- data.frame(predict(dmy, newdata = df))

glimpse(dat_transformed)
```
```{r}
colnames(dat_transformed)
```
```{r}
df1 <- dat_transformed
```

```{r}
colnames(df)
```

```{r}
df1$Severity.1 <- NULL
df1$Street.0S539.Summit.Ave <- NULL
df1$Side.L <- NULL
df1$City.Aberdeen <- NULL
df1$State.AL <- NULL
df1$Timezone.US.Central <- NULL
df1$Airport_Code.K04W <- NULL
df1$Wind_Direction.CALM <- NULL
df1$Weather_Condition.Blowing.Snow <- NULL
df1$Amenity.FALSE <- NULL
df1$Bump.FALSE <- NULL
df1$Crossing.FALSE <- NULL
df1$Give_Way.FALSE <- NULL
df1$Junction.FALSE <- NULL
df1$No_Exit.FALSE <- NULL
df1$Railway.FALSE <- NULL
df1$Roundabout.FALSE <- NULL
df1$Station.FALSE <- NULL
df1$Stop.FALSE <- NULL
df1$Traffic_Calming.FALSE <- NULL
df1$Traffic_Signal.FALSE <- NULL
df1$Sunrise_Sunset.Night <- NULL
df1$Civil_Twilight.Night <- NULL
df1$Nautical_Twilight.Night <- NULL
df1$Astronomical_Twilight.Night <- NULL
```

```{r}
df1$Severity.2 <- NULL
df1$Severity.3 <- NULL
df1$Severity.4 <- NULL 
```

```{r} 
#min-max
process <- preProcess(as.data.frame(df1), method = c("range"))
norm_scale <- predict(process, as.data.frame(df1))
```
```{r}
head(norm_scale)
```
```{r}
sev
```

```{r}
df2 <- norm_scale #encoded and scaled data
head(df2)
```
```{r}
df2$Severity <- sev
```

```{r}
str(df2$Severity)
```
```{r}
df2$Start_Time <- NULL
```

```{r}
library(caTools)
set.seed(100)
sample = sample.split(df2$Severity, SplitRatio = 0.75)
train = subset(df2, sample == T)
test = subset(df2, sample == F)
```

```{r}
library(caret)
```

```{r}
dim(train)
```

```{r}
dim(test)
```

```{r}
levels(df$Severity)
```
```{r}
svm1 <- svm(Severity~., data=train, 
          method="C-classification", kernal="radial", 
          gamma=0.1, cost=10)
```
```{r}
summary(svm1)
```
```{r}
prediction <- predict(svm1, test)
xtab <- table(test$Severity, prediction)
xtab
```
```{r}
#install.packages("performanceEstimation")
```
```{r}
library(e1071)
#accuracy
sum(diag(xtab))/sum(xtab)
```
```{r}
diag(xtab)/rowSums(xtab) # for each class per obs
```
```{r}
diag(xtab)/colSums(xtab) #For each class per prediction
```

